---
# title: "Week 2 Assignment"
title: "Report of NYPD Shooting Incident"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Data Source

The dataset *NYPD Shooting Incident Data (Historic)* was downloaded from [https://catalog.data.gov/dataset](https://catalog.data.gov/dataset). 


## Objective
The topic of this report is to show the whole picture of shooting incidents, including the trend, occurring time, occurring place, and profile of perpetrator and victims. At the end of the report, a SVM model is applied to forecast the victim's death or not.


## Import Library
```{r import_library, message=FALSE, warning=FALSE}

# R ver 4.1.2
# RStudio ver 2021.09.1
library(data.table)  # ver 1.14.2
library(dplyr)       # ver 1.0.7
library(lubridate)   # ver 1.8.0
library(stringr)     # ver 1.4.0
library(ggplot2)     # ver 3.3.5
library(knitr)       # ver 1.36
library(caret)       # ver 6.0-81
library(ROSE)        # ver 0.0-4
library(e1071)       # ver 1.7-1


# Remove scientific notation
options(scipen = 100)

```

## Clean Data

This dataset has 23,568 rows, and some columns contains na or blank values, like JURISDICTION_CODE and PERP_AGE_GROUP.

```{r clean_data}

data <- read.csv(paste0(getwd(), '/NYPD_Shooting_Incident_Data__Historic_.csv')) %>% 
  setDT() %>% 
  .[, ':='(X_COORD_CD = as.numeric(X_COORD_CD),
           Y_COORD_CD = as.numeric(Y_COORD_CD),
           JURISDICTION_CODE = as.factor(JURISDICTION_CODE),
           PRECINCT = as.factor(PRECINCT),
           STATISTICAL_MURDER_FLAG = ifelse(STATISTICAL_MURDER_FLAG=='true',
                                            1, 0))
    ] %>% 
  .[, ':='(INCIDENT = 1,
           YEAR = substr(OCCUR_DATE, 7, 11),
           MONTH = substr(OCCUR_DATE, 1, 2),
           DAY = substr(OCCUR_DATE, 4, 5),
           OCCUR_HOUR = substr(OCCUR_TIME, 1, 2))
    ] %>% 
  .[, ':='(YEAR = as.integer(YEAR),
           MONTH = as.integer(MONTH),
           DAY = as.integer(DAY),
           OCCUR_HOUR = as.integer(OCCUR_HOUR),
           YEAR_MONTH = paste0(YEAR, MONTH),
           OCCUR_DATE = ymd(paste0(YEAR, MONTH, DAY)))
    ]
```


## Trend of Incidents
First, I was curious about the trend of shooting incidents. The number of incidents decreased before 2020, however, increased dramatically in 2020.


```{r trend}

# 3.1 Trend of Incidents
line_chart_data <- copy(data) %>%
  .[, .(INCIDENT = sum(INCIDENT)), by = .(YEAR_MONTH)] %>%
  setorder(YEAR_MONTH)
  
# Set X axis
year_month <- line_chart_data$YEAR_MONTH
x_names = c()

for(i in 1:length(line_chart_data$YEAR_MONTH)){
  if(i %% 30 == 0 || i == 1){
    x_names[i] <- year_month[i]
  } else{
    x_names[i] <- ""
  }
}

# Plot
line_chart_data %>% 
  ggplot( aes(x=YEAR_MONTH, y=INCIDENT, group = 1)) +
  geom_line(size=0.3) +
  scale_x_discrete(name="YEAR_MONTH", 
                   labels=x_names)

```

## When The Shooting Incidents Occurred
It's obvious that most of shooting incidents occurred in night, and the peak was around 11 P.M.

```{r}
occur_hour <- copy(data) %>% 
  .[, .(INCIDENT = sum(INCIDENT)),
    by = .(OCCUR_HOUR)
    ] %>% 
  setorder(OCCUR_HOUR)

ggplot(data=occur_hour, aes(x = OCCUR_HOUR, y = INCIDENT)) + 
  geom_col(fill="#56B3F8")

```


## Where The Shooting Incidents Occurred

I have two plots to show where the shooting incidents occurred. The first one is a histogram by borough, and the second one is a scatter plot by longtitude and latitude.

```{r boro_hist}
boro_incident <- copy(data) %>% 
  .[, .(INCIDENT = sum(INCIDENT)), by = .(BORO)]

ggplot(data, aes(x=BORO, y=INCIDENT)) + 
  geom_col(fill="#56B3F8")
```


```{r map}
map_data <- copy(data) %>% 
  .[!is.na(JURISDICTION_CODE), ] %>% 
  .[, ':='(JURISDICTION = case_when(
    JURISDICTION_CODE == "0" ~ "Patrol",
    JURISDICTION_CODE == "1" ~ "Transit",
    JURISDICTION_CODE == "2" ~ "Housing"))
    ]

ggplot(map_data, aes(x=Longitude, y=Latitude, color=JURISDICTION)) + 
  geom_point(size=0.2)

```


## Profile of Perpetrator And Victim

In this section, I will show the features of perpetrator and victims. Through the heatmaps, around 50% of shooting incidents were related with black people.

### Profile of Perpetrator

```{r prep}

perp <- copy(data) %>% 
  .[!PERP_RACE %in% c("", "UNKNOWN") 
    & !PERP_AGE_GROUP %in% c("", "UNKNOWN")
    & PERP_RACE != "", 
    ] %>% 
  .[str_detect(PERP_AGE_GROUP, "<") 
    | str_detect(PERP_AGE_GROUP, "-")
    | str_detect(PERP_AGE_GROUP, "\\+"), ] %>% 
  .[, .(INCIDENT = sum(INCIDENT)),
    by = .(PERP_RACE, PERP_AGE_GROUP)
    ] %>% 
  .[, ':='(TOTAL = sum(INCIDENT))] %>% 
  .[, ':='(INCIDENT_RATIO = INCIDENT / TOTAL)]


ggplot(perp, aes(x = PERP_RACE, y = PERP_AGE_GROUP, fill = INCIDENT_RATIO)) +
  geom_tile(aes(fill = INCIDENT_RATIO)) +
  guides(x = guide_axis(angle = 15)) +
  theme(plot.margin=unit(c(0.5,0,0.5,3.5), "cm"))

```

### Profile of Victim

```{r victim}

# 4.2 Victim
victim <- copy(data) %>% 
  .[!VIC_RACE %in% c("", "UNKNOWN") 
    & !VIC_AGE_GROUP %in% c("", "UNKNOWN"), ] %>% 
  .[str_detect(VIC_AGE_GROUP, "<") 
    | str_detect(VIC_AGE_GROUP, "-")
    | str_detect(VIC_AGE_GROUP, "\\+"), ] %>% 
  .[, .(INCIDENT = sum(INCIDENT)),
    by = .(VIC_RACE, VIC_AGE_GROUP)
    ] %>% 
  .[, ':='(TOTAL = sum(INCIDENT))] %>% 
  .[, ':='(INCIDENT_RATIO = INCIDENT / TOTAL)]

ggplot(victim, aes(x = VIC_RACE, y = VIC_AGE_GROUP, fill = INCIDENT_RATIO)) +
  geom_tile(aes(fill = INCIDENT_RATIO)) +
  guides(x = guide_axis(angle = 15)) +
  theme(plot.margin=unit(c(0.5,0,0.5,3.5), "cm"))

```

### Common Factors of Perpetrator And Victim

```{r prep_victim}

# 4.2 Perpetrator And Victim
perp_vic_race <- copy(data) %>% 
  .[!VIC_RACE %in% c("", "UNKNOWN") 
    & !VIC_AGE_GROUP %in% c("", "UNKNOWN")
    & PERP_RACE != "", 
    ] %>% 
  .[str_detect(VIC_AGE_GROUP, "<") 
    | str_detect(VIC_AGE_GROUP, "-")
    | str_detect(VIC_AGE_GROUP, "\\+"), ] %>% 
  .[, .(INCIDENT = sum(INCIDENT)),
    by = .(PERP_RACE, VIC_RACE)
    ] %>% 
  .[, ':='(TOTAL = sum(INCIDENT))] %>% 
  .[, ':='(INCIDENT_RATIO = INCIDENT / TOTAL)]

ggplot(perp_vic_race, 
       aes(x = PERP_RACE, y = VIC_RACE, fill = INCIDENT_RATIO)) +
  geom_tile(aes(fill = INCIDENT_RATIO)) +
  guides(x = guide_axis(angle = 15)) +
  guides(x = guide_axis(angle = 15)) +
  theme(plot.margin=unit(c(0.5,0,0.5,0.5), "cm"))

```

## Prediction Murder

If we can predict a shooting incident will occur, and will it cause any person death, then NYPD can arrange more police officers to patrol at the specific time and place.

### Ratio of Murder
Before building a model, it is helpful to know the ratio of murder. Definitely, the ratio is 19%, which means the data is imbalanced.

```{r death_ratio}

murder_flag <- copy(data) %>% 
  .[, .(INCIDENT = sum(INCIDENT))
    , by = .(STATISTICAL_MURDER_FLAG)
  ] %>% 
  .[, ':='(STATISTICAL_MURDER_FLAG = ifelse(STATISTICAL_MURDER_FLAG==0, 
                                            'N', 'Y'))]

ggplot(murder_flag, aes(x=STATISTICAL_MURDER_FLAG, 
                        y=INCIDENT)) + 
  geom_col(fill="#56B3F8")

```


### Prediction

```{r prediction}

# SVM Model
model_data <- data %>%
  .[, .(BORO, JURISDICTION_CODE, OCCUR_HOUR,
        PERP_RACE, VIC_RACE, STATISTICAL_MURDER_FLAG,
        Latitude, Longitude, DAY, MONTH)
    ]

dummy_data <- copy(model_data) %>% 
  .[PERP_RACE != "" & !is.na(JURISDICTION_CODE)
    & !is.na(OCCUR_HOUR), ]

dmy <- dummyVars(" ~ .", data = dummy_data)
dummy_data <- data.frame(predict(dmy, newdata=dummy_data))
dummy_data <- dummy_data %>% setDT()


# Standardization Data
norm_data <- preProcess(dummy_data, method=c("range"))
norm_data <- predict(norm_data, dummy_data)

# Split train data and test data
train_index <- sample(seq_len(nrow(norm_data)),
                      size = 0.95 * nrow(norm_data))

train <- norm_data[train_index, ]
test <- norm_data[-train_index, ]
test_var <- test %>% 
  .[, -c('STATISTICAL_MURDER_FLAG'), with=FALSE]

test_true <- as.factor(test$STATISTICAL_MURDER_FLAG) 

# Handel Imbalanced Data
table(train$STATISTICAL_MURDER_FLAG)
balenced_train <- ovun.sample(STATISTICAL_MURDER_FLAG ~ .,
                             data = train,
                             method = "over")$data

balenced_train <- as.data.frame(balenced_train) %>% 
  setDT()

table(balenced_train$STATISTICAL_MURDER_FLAG)


# Build a SVM model
model <- svm(STATISTICAL_MURDER_FLAG~., data = balenced_train,
             type = 'C-classification',
             kernel = "linear", scale = FALSE)

pred <- predict(model, test_var, na.action = na.pass)
pred <- predict(model, test_var)

confusionMatrix(pred, test_true)


```


## Conclusion And Bias Detection

According to the report, the shooting incidents usually happened in specific borough by black people, but we couldn't say black people tend to commit a shooting incident. It is obvious that the data source was incomplete and had bias. If we could add more data into this project, like economy, salary, and living quality, the conclusion would be more reasonable and convincible.

<br>
<br>
<br>
<br>

